Legacy Data Confounds Modern Genomics Studies

### 2019-08-08 LatexManuscript/MBE/legacy_mbe.tex
moved things to one folder for mbe upload

%%% PREAMBLE 
\documentclass[9pt,article]{template}

\usepackage[export]{adjustbox}
\usepackage{lscape}
\usepackage{afterpage}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{ulem}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

\newcommand{\sgcomment}[1]{\textcolor{blue}{SG: #1}}
\newcommand{\luke}[1]{\textcolor{blue}{Luke: #1}}
\newcommand{\todo}[1]{\textcolor{blue}{*#1*}}
\newcommand{\alex}[1]{\textcolor{red}{Alex: #1}}
\newcommand{\logit}{\operatorname{logit}}
%\newcommand{\qq}{m}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Legacy Data Confounds Genomics Studies}

\author[1,2]{Luke Anderson-Trocm\'e}
\author[1,2]{Rick Farouni}
\author[1,2]{Mathieu Bourgey}
\author[3]{Yoichiro Kamatani}
\author[3]{Koichiro Higasa}
\author[4,5]{Jeong-Sun Seo}
\author[4]{Changhoon Kim}
\author[3]{Fumihiko Matsuda}
\author[1,2]{Simon Gravel}

\affil[1]{Department of Human Genetics, McGill University, Montreal, QC H3A 0G1, Canada}
\affil[2]{McGill University and Genome Quebec Innovation Centre, Montreal, QC H3A 0G1, Canada}
\affil[3]{Center for Genomic Medicine, Graduate School of Medicine, Kyoto University, Kyoto 606-8501, Japan}
\affil[4]{Bioinformatics Institute, Macrogen Inc., Seoul, 08511, Republic of Korea}
\affil[5]{Precision Medicine Center, Seoul National University Bundang Hospital, Seongnam, 13605, Republic of Korea}
\corr{simon.gravel@mcgill.ca}{SG}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\maketitle
\begin{abstract}
Recent reports have identified differences in the mutational spectra across human populations. While some of these reports have been replicated in other cohorts, most have been reported only in the 1000 Genomes Project (1kGP) data. While investigating an intriguing putative population stratification within the Japanese population, we identified a previously unreported batch effect leading to spurious mutation calls in the 1kGP data and to the apparent population stratification. Because the 1kGP data is used extensively, we find that the batch effects also lead to incorrect imputation by leading imputation servers and a small number of suspicious GWAS associations. Lower-quality data from the early phases of the 1kGP thus continues to contaminate modern studies in hidden ways. It may be time to retire or upgrade such legacy sequencing data.
\vspace{5mm}

\textbf{Key words} : Batch Effect, Mutational Signature, Statistical Genetics, Population Genetics, Reference Cohorts, Imputation
\end{abstract}

\section{Introduction}

@@

\bibliography{Legacy}

>>>>

\documentclass[nogrid]{MBE}%

\usepackage{url}
%\usepackage[T1]{fontenc}
\usepackage[export]{adjustbox}
\usepackage{lscape}
\usepackage{afterpage}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{ulem}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
\usepackage{xcolor}

\jshort{mst}

\volname{}

\jvolume{0}

\jvol{}

\jissue{0}

\pubyear{2019}

\mstype{Article}

\artid{012}

\newcommand{\sgcomment}[1]{\textcolor{blue}{SG: #1}}
\newcommand{\luke}[1]{\textcolor{blue}{Luke: #1}}
\newcommand{\todo}[1]{\textcolor{blue}{*#1*}}
\newcommand{\alex}[1]{\textcolor{red}{Alex: #1}}
\newcommand{\logit}{\operatorname{logit}}

\title{ {{L}}egacy Data Confounds Genomics Studies}

\author[Anderson-Trocm\'e et al.]{
{{L}}uke \surname{Anderson-Trocm\'e},$^{\ast,1,2}$ 
Rick Farouni,$^{1,2}$ 
Mathieu Bourgey,$^{1,2}$ 
Yoichiro Kamatani,$^{3}$ 
Koichiro Higasa,$^{3}$ 
Jeong-Sun Seo,$^{4,5}$ 
Changhoon Kim$^{4}$ 
Fumihiko Matsuda$^{3}$ 
and Simon Gravel$^{1,2}$}

\address{$^{1}$Department of Human Genetics, McGill University, Montreal, QC H3A 0G1, Canada\\
$^{2}$McGill University and Genome Quebec Innovation Centre, Montreal, QC H3A 0G1, Canada\\ 
$^{3}$Center for Genomic Medicine, Graduate School of Medicine, Kyoto University, Kyoto 606-8501, Japan\\ 
$^{4}$Bioinformatics Institute, Macrogen Inc., Seoul, 08511, Republic of Korea\\ 
$^{5}$Precision Medicine Center, Seoul National University Bundang Hospital, Seongnam, 13605, Republic of Korea}

\history{Received 3 May 2019; reviews returned 27 May 2019}

\coresp{E-mail: simon.gravel@mcgill.ca}

\editor{Sara Tinsley}

\abstract{Recent reports have identified differences in the mutational spectra across human populations. While some of these reports have been replicated in other cohorts, most have been reported only in the 1000 Genomes Project (1kGP) data. While investigating an intriguing putative population stratification within the Japanese population, we identified a previously unreported batch effect leading to spurious mutation calls in the 1kGP data and to the apparent population stratification. Because the 1kGP data is used extensively, we find that the batch effects also lead to incorrect imputation by leading imputation servers and a small number of suspicious GWAS associations. Lower-quality data from the early phases of the 1kGP thus continues to contaminate modern studies in hidden ways. It may be time to retire or upgrade such legacy sequencing data.}

\keyword{Batch Effect, Mutational Signature, Statistical Genetics, Population Genetics, Reference Cohorts, Imputation}

@@

\maketitle

\section{{Introduction}\label{sec:Intro}}

@@

\bibliographystyle{natbib}%%%%natbib.sty
\bibliography{legacy_mbe}%%%Legacy.bib

### 2019-08-08 PlottingScripts/ProcessPlotData.Rmd
moved things to one folder for mbe upload

#Get P-Values and significant sites
Reg_mask$plog10 <- -log10(pchisq(Reg_mask$deviance, 1, 
                            lower.tail = FALSE))
Reg_mask<-Reg_mask[which(Reg_mask$plog10 >= 6),]
#Count number of populations where a variant is significant
NumPop <- as.data.frame(table(Reg_mask$Chr,Reg_mask$Pos))
names(NumPop) <- c('Chr','Pos', 'SigHits')
NumPop$Chr<-as.numeric(as.character(NumPop$Chr)) 
NumPop$Pos<-as.numeric(as.character(NumPop$Pos))

>>>>

#Get P-Values and significant sites
Reg_mask$plog10 <- -log10(pchisq(Reg_mask$deviance, 1, 
                            lower.tail = FALSE))
Reg_mask<-Reg_mask[which(Reg_mask$plog10 >= 6),]
#Count number of populations where a variant is significant
NumPop <- as.data.frame(table(Reg_mask$Chr,Reg_mask$Pos))
names(NumPop) <- c('Chr','Pos', 'SigHits')
NumPop$Chr<-as.numeric(as.character(NumPop$Chr)) 
NumPop$Pos<-as.numeric(as.character(NumPop$Pos))
all_sites <- merge(Reg_mask,
                   NumPop, 
                   by = c('Chr','Pos')) %>%
  group_by(Population, strict) %>%
  summarize(total_count = n())
all_sites[which(all_sites$Population%in% phase1),]$Population <- 
  paste0('*',all_sites[which(all_sites$Population%in% phase1),]$Population)
